# R Code {#appendix-a}

## Implementations of Model Error Methods {#noisemaker-code}

The implementations of the model error methods used are also available bundled as an R package, *noisemaker*, which can be downloaded from [https://www.github.com/JustinKracht/noisemaker](https://www.github.com/JustinKracht/noisemaker).

```{r find-code-files, echo = FALSE, include = FALSE}
code_files <- list.files(here::here("code/noisemaker"), full.names = TRUE)
code_files <- code_files[stringr::str_ends(code_files, ".R")]

# Function to read code files and include them in R Markdown
read_files <- function(files) {
  unlist(lapply(files, xfun::read_utf8))
}
```

\small

```{r write-noisemaker-code, eval = FALSE, code = read_files(code_files), echo = TRUE}

```

## Main Simulation Code {#main-simulation}

```{r find-simulation-code-files}
simulation_code_files <- list.files(here::here("code/model-error-simulation"), 
                                    full.names = TRUE)
simulation_code_files <- simulation_code_files[stringr::str_ends(simulation_code_files, ".R")]
```

\small

```{r write-simulation-code, eval = FALSE, code = read_files(simulation_code_files), echo = TRUE}

```

## Secondary Simulation: Check Effect of $\lambda$ Values {#effect-of-lambda}

\small

```{r check-lambda-values, eval = FALSE, echo = TRUE}
# Set the number of reps
reps <- 200

# Create a matrix of fully-crossed conditions
factors <- unique(results_matrix$factors)
items_per_factor <- 15
loading <- c(.4)
target_rmsea <- c(0.090)
penalty <- c(0, .1, 1, 1e1, 1e2, 1e3, 1e4, 1e5, 1e6)

conditions_matrix <- expand.grid(
  factors = factors,
  items_per_factor = items_per_factor,
  loading = loading,
  target_rmsea = target_rmsea
)


# Function to check whether a W matrix has more than two factor loadings greater
# than 0.3 for any minor factor.
check_constraints <- function(W) {
  any(max(apply(abs(W) >= .3, 2, sum)) > 2)
}

# For each condition, generate a population correlation matrix without model
# error and then generate 200 population correlations with model error using
# the TKL (RMSEA) method
set.seed(666)
constraint_violations <- map_dfr(
  .x = 1:nrow(conditions_matrix), 
  .f = function(condition) {
    
    cat("\nWorking on condition", condition, "of", nrow(conditions_matrix))
    
    # Generate population correlation matrix without model error
    mod <- simFA(
      Model = list(NFac = conditions_matrix$factors[condition],
                   NItemPerFac = conditions_matrix$items_per_factor[condition],
                   Model = "orthogonal"),
      Loadings = list(FacLoadDist = "fixed",
                      FacLoadRange = conditions_matrix$loading[condition])
    )
    
    # Generate 200 population correlation matrices with model error
    pro_map_dfr(.x = 1:reps,
                .f = function(x, target_rmsea) {
                  map_dfr(.x = penalty, 
                          .f = function(penalty, mod, target_rmsea, seed) {
                            set.seed(seed)
                            sol <- noisemaker(mod = mod, 
                                              method = "TKL", 
                                              target_rmsea = target_rmsea,
                                              target_cfi = NULL,
                                              tkl_ctrl = list(penalty = penalty,
                                                              NWmaxLoading = 2,
                                                              WmaxLoading = .3))
                            w_constraints_violated <- check_constraints(sol$W)
                            
                            c(penalty = penalty, 
                              constraints_violated = w_constraints_violated)
                          },
                          mod = mod,
                          target_rmsea = target_rmsea,
                          seed = sample(1e6, 1))
                }, target_rmsea = conditions_matrix$target_rmsea[condition])
  }
)

# Bind the conditions matrix to the results to indicate which result belongs to
# which condition
constraint_violations <- bind_cols(
  conditions_matrix[rep(1:nrow(conditions_matrix), 
                        each = length(penalty) * reps),],
  constraint_violations
)

# Plot the results
constraint_violations %>%
  mutate(factors = as.factor(factors),
         penalty = factor(penalty, 
                          labels = c("0", "0.1", "1", "10", 
                                     "100", "1,000", "10,000",
                                     "100,000", "1,000,000"))) %>%
  group_by(factors, penalty) %>%
  summarise(percent = mean(constraints_violated, na.rm = TRUE)) %>%
  ggplot(aes(x = penalty, y = percent, color = factors, shape = factors, 
             linetype = factors, group = factors)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_brewer(palette = "Dark2", type = "qual") +
  labs(y = "Cases with Violtated W Constraints",
       x = latex2exp::TeX("$\\lambda$"),
       color = "Factors", shape = "Factors", linetype = "Factors") +
  theme_bw() +
  theme(legend.position = "bottom")

# Save the plot
ggsave(filename = here("img/penalty_values.png"),
       dpi = 320,
       height = 4,
       width = 6)
```

## Secondary Simulation: Check Whether $\textrm{TKL}_{\textrm{CFI}}$ Requires A Penalty {#tkl-cfi-penalty}

\small

```{r check-whether-tkl-cfi-needs-penalty, eval = FALSE, echo=TRUE}
# Check whether the TKL (CFI) method leads to solutions that violate the W 
# constraints when no penalty is applied
set.seed(42)

# Create a population correlation matrix without model error (Omega)
mod <- simFA(Model = list(NFac = 10,
                          NItemPerFac = 15,
                          Model = "orthogonal"),
             Loadings = list(FacLoadDist = "fixed",
                             FacLoadRange = .4))

# Function to check whether a W matrix has more than two factor loadings greater
# than 0.3 for any minor factor.
check_constraints <- function(W) {
  any(max(apply(abs(W) >= .3, 2, sum)) > 2)
}

# Generate 200 population correlation matrices with model error (Sigma) using 
# the TKL (CFI) method without a penalty, then check how often the constraints
# on W were violated
constraints_violated_vec <- pbmclapply(
  X = 1:200,
  FUN = function(x) {
    sol <- noisemaker(mod = mod, 
                      method = "TKL", 
                      target_rmsea = 0.09,
                      target_cfi = 0.9,
                      tkl_ctrl = list(penalty = 0))
    check_constraints(sol$W)
  },
  mc.cores = 4
)

# What percent of cases had violated W constraints?
mean(constraints_violated_vec, na.rm = TRUE)
```

## Secondary Simulation: Check Recovery of Model Fit Indices with Known-Possible Values {#check-model-fit-recovery-code}

```{r find-secondary-simulation-code-files}
secondary_simulation_code_files <- list.files(here::here("code/misc"), 
                                    full.names = TRUE)
secondary_simulation_code_files <- secondary_simulation_code_files[stringr::str_ends(secondary_simulation_code_files, ".R")]
```

\small

```{r write-secondary-simulation-code, eval = FALSE, code = read_files(secondary_simulation_code_files), echo = TRUE}

```



